F-stats vs PCA

### 2021-03-19 PNAS-template-main.tex
cleaner

@@

\title{The Geometry of Admixture}

@@

@@

As $(X_{ai}-X_{xi})(X_{bi}-X_{xi}) = (X_{ai}+\mu_i -X_{xi} - \mu_i)(X_{bi} + \mu_i-X_{xi}_i-\mu_i) = (Y_{ai}-Y_{xi})(Y_{bi}-Y_{xi})$, we can rewrite $F_3$ as

@@

\subsection*{When is F3-negative?}
Given two samples $A$ and $B$, for which sample $X$ will we detect admixture?
As 
\begin{equation}
    2 F_3(A, B,X) = \nsq{A-X} + \nsq{B-X} - \nsq{A-B},
\end{equation}
it follows from the Pythagorean theorem that $F_3$ will be zero if  if $X$ lies on the $K$-dimensional sphere with diameter $$\norm{A,B}$, negative if it lies inside and positive outside (FIGURE).

>>>>

\include{amsmath}

@@

\title{F-stats vs PCA}

@@

$F_3$ and $F_4$ can be written as functions of $F_2$ \cite{peter2016}: 
\begin{align}
F_3(P_A, P_B; P_X) &= F_2(P_A, P_X) + F_2(P_B, P_X) - F_2(P_A, P_B)\\
F_4(P_A, P_B; P_C, P_D) &= F_2(P_A, P_C) + F_2(P_B, P_D) - F_2(P_A, P_C) - F_2(P_B, P_C)
\end{align}

In many applications, we will have a large set of $n$ samples that we wish to analyze. One way to proceed is to calculate all $F_2$ statistics, which will be a sufficient summary to calculate all $F_3$ and $F_4$ statistics from this data.

$F_2$ can also be written as 
$$\mathbb{E}F_2 \propto 2T_{12} - T_{11} - T_{22}$$

Similarly, we may want to calculate a PCA:

@@

As $(X_{ai}-X_{xi})(X_{bi}-X_{xi}) = (X_{ai}+\mu_i -X_{xi} - \mu_i)(X_{bi} + \mu_i-X_{xi}-\mu_i) = (Y_{ai}-Y_{xi})(Y_{bi}-Y_{xi})$, we can rewrite $F_3$ as

@@

\subsection*{When is F3-negative?}
Given two samples $A$ and $B$, for which sample $X$ will we detect admixture?
As 
\begin{equation}
    2 F_3(A, B,X) = \nsq{A-X} + \nsq{B-X} - \nsq{A-B},
\end{equation}
it follows from the Pythagorean theorem that $F_3$ will be zero if  if $X$ lies on the $K$-dimensional sphere with diameter $\norm{A,B}$, negative if it lies inside and positive outside (FIGURE).
